### 2021-03-10 paper/rebuttal/rebuttal.tex

\bigskip
\thispagestyle{empty}
\vspace{-0.4cm}
\noindent Dear Dr. Sutherland and Referees,\\
\vspace{-0.25cm}

We thank all the referees for their thoughtful and helpful comments and critiques.
We acknowledge that our initial submission had a substantial gap regarding the choice of the \textit{word2vec} model, its connection with the gravity model of mobility, and its evaluation.
While reflecting on these comments and our manuscript, we have made a theoretical breakthrough that addresses the fundamental issues;
namely, we have discovered \emph{a mathematical equivalence between the \textit{word2vec} (skip-gram with negative sampling) model and the gravity model.}

Furthermore, our subsequent experiments show that our approach's performance is robust, exceeding many other baselines, even direct optimization of the gravity model.
Therefore, with these changes, we are confident that our paper is substantially stronger with a qualitative leap.
Thus, we would like to request a reconsideration of our manuscript.
Again, we are incredibly appreciative of these comments, which provided the insight and motivation needed to make these breakthroughs.

We have developed the mathematical and theoretical basis of our approach for embedding mobility, and made additional updates to our experiments and manuscript, addressing the three main issues raised by referees:
\vspace{-0.1cm}
\begin{enumerate}
	\itemsep0em
	\item The lack of theoretical justification and grounding for the application of \textit{word2vec}
	\item The lack of simpler baselines to understand the performance of \textit{word2vec}
	\item The lack of sufficient grounding in the literature of neural embeddings and mobility
\end{enumerate}
\vspace{-0.1cm}
We address the first point by presenting the mathematical equivalence between the Gravity Model of Mobility and the skip-gram negative sampling \textit{word2vec} model.
Namely, we demonstrate that \textit{word2vec}'s objective function is equivalent to the gravity model and thus \textit{word2vec} learns gravity-like relationships between entities, whether it be language or locations.
This development constitutes a major breakthrough in our understanding of neural embeddings and provides a theoretical justification for their application to mobility.

@@

\vspace{-0.3cm}
\paragraph{Laplacian Eigenmap:} Similarly, we consider Laplacian Eigenmap embeddings~\autocite{belkin2003laplacian} as a baseline, which is one of the most fundamental approaches for graph embedding. First, we construct the co-occurrence matrix of $N$ organizations, $A_{ij}$ given by the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory and degree matrix $D$ which is the diagonal matrix for which $D_{ii}=\sum_j {A_{ij}}$. Then we construct graph Lapalcian matrix $L = D-A$  and apply truncated singular vector decomposition in the matrix $L$ with $d=300$.

\vspace{-0.3cm}
\paragraph{Levy's  symmetric SVD \textit{word2vec} distance:} In addition to these suggested baselines, we also consider distances derived from embeddings learned through matrix factorization following the procedure outlined by Levy \& Goldberg~\autocite{levy2014neural}. Based on idea of \textit{word2vec} is implicit matrix factorization, Levy proposed symmetric SVD \textit{word2vec} embedding, which should directly compute the embedding that \textit{word2vec} only attempts to efficiently approximate. First, we construct the matrix of $N$ organization
\begin{equation}
	M_{ij} = \log\left(\frac{N(i,j) |D|}{N(i) N(j)}\right) - \log(k),
\end{equation}
where $N(i,j)$ is the number of times the location pair $(i,j)$ appears given the window size $w$ in the total corpus $D$, $N(i)=\sum_{j=N}^i N(i,j)$ as the number of items $i$ occurred given the window size $w$ in $D$, and k is the number of negative samples. We used $w=1$ and $k=5$ which is same setting in the our main result. Then, we factorized matrix $M$ with truncated singular vector decomposition in the matrix with $d=300$ into $U_d \Sigma_d V_d$, and used the embedding vector as $U_d\sqrt{\Sigma_d}$.

\vspace{-0.3cm}
\paragraph{Direct optimization of gravity model:} Finally, we create an embedding by optimizing for the gravity model directly.
Specifically, we construct a gravity matrix of $N$ organizations, where each cell of the matrix is calculated as $T_{ij}/m_i m_j$, where $T_{ij}$ is the co-occurrence of organizations $i$ and $j$ in the same affiliation trajectory, and $m_i$ and $m_j$ are the organizations' populations, defined here as the mean annual member of unique mobile and non-mobile authors affiliated. We then embed the gravity matrix using two approaches: a truncated singular-vector decomposition (SVD), and multidimensional scaling (MDS), which embeds each location in an N-dimensional space such that pairwise distances are preserved as well as possible.
Typically, MDS uses euclidean distance to measure the distance between vectors.

@@

	\begin{itemize}
		\item Liu, Xin, Yong Liu, and Xiaoli Li. ``Exploring the Context of Locations for Personalized Location Recommendations.'' IJCAI. 2016.
		\item Feng, Shanshan, et al. ``POI2Vec: geographical latent representation for predicting future visitors.'' AAAI. 2017.
		\item Yao, Zijun, et al. ``Representing Urban Functions through Zone Embedding with Human Mobility Patterns.'' IJCAI. 2018.
		\item Cao, Hancheng, et al. ``Habit2vec: Trajectory semantic embedding for living pattern recognition in population.'' IEEE Transactions on Mobile Computing 19.5 (2019): 1096-1108.
	\end{itemize}

	However, the novelty of this work is quite narrow, as there are several prior works that develop trajectory-based embeddings of human mobility based on the word2vec algorithm. For example, the following references all propose word2vec-inspired embeddings of human mobility data:

	Notably, the location embedding methodology in Liu et al (2016) (see Section 3.1) is essentially the same as the approach used in the submitted manuscript, and the work by Liu et al (2016) is well-known with over 100 citations.
}

\response{Thank you for pointing out the missing references!
	We have carefully read through each, and learned a lot in the process.
	They have now been incorporated into our manuscript's introduction.
	Although there are existing works on trajectory-based mobility embedding, we argue that our new result establishing the mathematical equivalence of \textit{word2vec} and the gravity model, as well as our extensive evaluation with many baselines represents a novel and substantial contribution.
}

@@

%
% R2 Table
%
\clearpage

\begin{landscape}

	\providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
	\providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
	\providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
	\providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

	\begin{table}[ht]
		\caption{
			\textbf{Correlation between flux and distance over metrics, experimental parameters}.
			Each cell corresponds to the correlation (measured with $R^{2}$) between the real-world flux between scientific organizations, comparing the embedding distance against baseline metrics, and across subsets of mobility data and by definitions of organization population.
			The asterisk denotes the top-performing distance metric by column.
			Distance metrics are ordered from highest $R^{2}$ to lowest, based on global mobility with organization population defined using all mobile and non-mobile authors.
			``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
			population is defined in the same way for ``Mobile only'', except only using unique mobile researchers;
			``Raw freq'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding.
			Embedding distance, measured as the cosine distance between embedding vectors, explains more of the flux than baselines in nearly every case, except using raw frequency population and domestic and international mobility, where direct optimization of the gravity model works better, as well as Levy's factorization~\autocite{levy2014neural} for domestic and international only mobility.
		}
		\label{supp:table:r2_table}
		\begin{centerbox}
			\begin{threeparttable}
				\small
				\setlength{\tabcolsep}{0pt}
				\begin{tabular}{l l l l l l l l l l}

@@

\end{landscape} % End r2 table page

%
% RMSE TABLE --- Power Law
%
\clearpage
\begin{landscape}

	\providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
	\providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
	\providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
	\providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

	\begin{table}[ht]
		\caption{
			\textbf{Prediction error between actual and predicted mobility with power-law gravity model, by metrics and experimental parameters}.
			Each cell corresponds to the prediction error (measured with root mean squared error) between the real-world flux between scientific organizations and predictions made with the power-law form of the gravity model of mobility.
			The prediction error is shown for the embedding distance compared against baselines, subsets of mobility data, and definitions of organization population.
			The asterisk denotes the top-performing distance metric by column (lowest prediction error).
			Distance metrics are ordered from lowest prediction error to highest, based on global mobility with organization population defined using all mobile and non-mobile authors.
			``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
			population is defined in the same way for ``Mobile only'', except only using unique mobile researchers;
			``Raw freq'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding.
			Embedding distance, measured as the cosine distance between embedding vectors, results in better predictions for global mobility with ``All'; and ``Mobile only'' definitions of population, though direct optimization of the gravity model with MDS performs better when using raw frequencies to measure organization population, and Levy's factorization~\autocite{levy2014neural} perform better here with domestic and international only mobility.
		}
		\label{supp:table:rmse_power_table}
		\begin{centerbox}
			\begin{threeparttable}
				\small
				\setlength{\tabcolsep}{0pt}
				\begin{tabular}{l l l l l l l l l l}
